### 2011-11-13 doc/theses/vitasek/chapter-2-research.tex
Thesis: ready to send to reviewers.

\chapter{Research}
\label{chapter-research}

@@

To the best of our knowledge, there are no other articles dealing directly with this problem.\\

Fajt in \cite{fajt} introduces several algorithms to find XML keys (see \cite{keX}) in existing XML documents. Simple keys found this way can be translated to an equivalent ID attribute definition. The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}. This opens a new line of research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to ID attributes.\\

>>>>

\chapter{Related Work}
\label{chapter-research}

@@

To the best of our knowledge, there are no other articles dealing with this problem.\\

Fajt in \cite{fajt} lists several algorithms to find XML keys (see \cite{keX}) in existing XML documents and introduces some new ones. XML keys found this way, being quite different from \texttt{ID} attributes, can under some circumstances (they have to be simple) be translated to an equivalent \texttt{ID} attribute definition. The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}. This opens a new line of possible research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to \texttt{ID} attributes.

In our work we find the \texttt{ID} attributes directly. Even though we can convert them to XML keys by the process mentioned above, we are unable to find more complex keys this way.\\

### 2011-11-13 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.
Thesis: ready to send to reviewers.

In this chapter we introduce a new approach to finding maximum ID sets. First, we transform the problem formulation to maximum weighted IS problem formulation. Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
We will continue by applying heuristical approaches to improve the performance of the process.

@@

GLPK \cite{glpk} is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} (see \cite{land60a} for details) for MIP. An advantage of using \textit{Branch \& Bound} is that while traversing the \textit{search tree} it finds intermittent, sub-optimal solutions. It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\

@@

\subsection{Example}

@@

First three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$. The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.

@@

It is obvious that this approach works and for any possible input we can let GLPK find the optimal solution. However, sometimes it takes too long to find the optimum (see e.g. Section \ref{section-glpk-comparison}), we should try to improve this process.

@@

\begin{define}[Heuristic]
	Is an approach to problem solving based on prior experience, educated guess or common knowledge.
\end{define}

\begin{define}[Heuristic Algorithm]
	Is an algorithm that, in a reasonably short time, will generate a good, maybe even optimal solution to an optimization problem. However, it won’t provide any formal guarantee about its quality.
\end{define}

@@

An example of a heuristic is the commonly used approach of \textit{trial and error}: after a failed attempt, change something and try again. We will see many more heuristics later in this chapter.

@@

\begin{define}[Solution Pool, Incumbent Solution]
	\textit{Solution pool} (sometimes called pool of feasible solutions or feasible pool) is simply a set of solutions of different qualities that were found in one of the stages of the metaheuristic. The solution(s) with the highest quality is (are) called the \textit{incumbent} solution(s).
\end{define}

@@

First we take the list of candidate AMs and ask a \textit{construction heuristic} (see Section \ref{section-mip-chs}) to provide us with a pool of solutions. Then, in a loop, we use this pool as input for \textit{improvement heuristics} (see Section \ref{section-mip-ihs}) and in turns ask them to improve it. All the time we check whether some \textit{termination criteria} are met, and if it is so, we terminate the metaheuristic. The incumbent solution from the last pool is then declared the Output ID set.\\

@@

@@

\begin{algorithm}
\caption{\heu{FIDAX} CH}
\label{listing-ch-fidax}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets C$ sorted by decreasing size
\STATE Compute the weight $w(m)$ of each $m$ in $C$
\FORALL{$t$ in $\Sigma^E$}
  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
  \STATE Remove from $C'$ all mappings of type $t$ except $m$
\ENDFOR
\FORALL{$m$ in $C'$}
  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
  \IF{$w(m) > \sum_{p \in S} w(p)$}
    \STATE remove all $p \in S$ from $C'$
  \ELSE
    \STATE remove $m$ from $C'$
  \ENDIF
\ENDFOR
\RETURN $C'$
\end{algorithmic}
\end{algorithm}

@@

This trivial heuristics sorts all candidate AMs by their decreasing weights (Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a hungry heuristic.

@@

This construction heuristic will be called \heu{Glpk} from now on, for it is basically a time-constrained run of GLPK. Recall that GLPK uses a the \textit{Branch \& Bound} algorithm that produces feasible solutions even before the optimum is found. Limiting the run time gives us the best solution found so far, which means this is a construction heuristic.

@@

\heu{Mutation} is based on the following idea. We assume that an incumbent solution may already contain some AMs belonging to the optimal solution. We will take a random guess and fix some of these AMs: add new constraints to the MIP formulation fixing values of the respective variables to 1.

@@

First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set. This is because the specification of DTD/XSD does not allow an attribute to be ID and IDREF (IDREFS) at the same time. Let us denominate these mappings as \textit{IDREF candidates} (obviously different from \textit{candidate AMs}).

>>>>

In this chapter we introduce a new approach to finding maximum ID sets. First, we transform the problem formulation to maximum weighted IS problem formulation. Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK \cite{glpk}. \nomenclature{GLPK}{GNU Linear Programming Kit}
We will continue by applying heuristical approaches to improve the performance of the process.

@@

GLPK is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} for MIP.

\paragraph{\textit{Branch \& Bound}} This is an optimization method where the search space is systematically divided into smaller sub-spaces. This is the \textit{branching} component, and a so-called \textit{search tree} is built this way. Then, the sub-problems are recursively solved and whole branches of the search tree are discarded when it becomes obvious that the solution does not lie there. This is the \textit{bounding} component. See \cite{land60a} for detailed description.\\

An advantage of using \textit{Branch \& Bound} is that while traversing the search tree it finds intermittent, sub-optimal solutions. It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\

@@

\subsection*{Example}

@@

The first three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$. The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.

@@

It is obvious that this approach works and for any possible input we can let GLPK find the optimal solution. However, sometimes it takes too long to find the optimum (see e.g. Section \ref{section-glpk-comparison}), hence, the aim of this work is to improve this process.

@@

\begin{define}[Heuristic]
	A \textit{heuristic} is an approach to problem solving based on prior experience, educated guess or common knowledge.
\end{define}

\begin{define}[Heuristic Algorithm]
	A \textit{heuristic algorithm} is one that, in a reasonably short time, generates a good, maybe even optimal solution to an optimization problem. However, it won’t provide any formal guarantee about its quality.
\end{define}

@@

An example of a heuristic is the commonly used approach of \textit{trial and error}: after a failed attempt, change some parameter and try again. We will see many more heuristics later in this chapter.

@@

\begin{define}[Solution Pool, Incumbent Solution]
	\textit{Solution pool} (sometimes called pool of feasible solutions or feasible pool) is a set of solutions of different qualities that were found in one of the stages of the metaheuristic. The solution(s) with the highest quality is (are) called the \textit{incumbent} solution(s).
\end{define}

@@

First we take the list of candidate AMs and ask a \textit{construction heuristic} (see Section \ref{section-mip-chs}) to provide us with a pool of solutions. Then, in a loop, we use this pool as input for \textit{improvement heuristics} (see Section \ref{section-mip-ihs}) and in turns ask them to improve it. All the time we check whether \textit{termination criteria} are met, and if so, we terminate the metaheuristic. The incumbent solution from the last pool is then declared the Output ID set.\\

@@

The algorithm works in two steps. First, all candidate AMs are grouped according to their types, and for each type the AM with the highest weight is selected. Second, all these AMs (now called $C'$) are traversed in order of their decreasing size. For each AM $m$, a set $S$ of all conflicting AMs from $C'$ is found and weights of both $m$ and $S$ are calculated. Then the weights are compared and either $m$ or $S$ is removed from $C'$.

@@

\begin{algorithm}
\caption{\heu{FIDAX} CH}
\label{listing-ch-fidax}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets C$ sorted by decreasing size
\STATE Compute the weight $w(m)$ of each $m$ in $C$
\FORALL{$t$ in $\Sigma^E$}
  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
  \STATE Remove from $C'$ all mappings of type $t$ except $m$
\ENDFOR
\FORALL{$m$ in $C'$}
  \STATE $S \gets$ all mappings in $C'$ whose images intersect $\iota(m)$
  \IF{$w(m) > \sum_{p \in S} w(p)$}
    \STATE remove all $p \in S$ from $C'$
  \ELSE
    \STATE remove $m$ from $C'$
  \ENDIF
\ENDFOR
\RETURN $C'$
\end{algorithmic}
\end{algorithm}

@@

This trivial heuristics sorts all candidate AMs by their decreasing weights (see Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a hungry heuristic.

@@

This construction heuristic will be called \heu{Glpk} from now on. It is basically a time-constrained run of GLPK. Recall that GLPK uses a the \textit{Branch \& Bound} algorithm that produces feasible solutions even before the optimum is found. Limiting the run time gives us the best solution found so far, which means this is a construction heuristic.

@@

\heu{Mutation} is based on the following idea. We assume that an incumbent solution may already contain some AMs belonging to the optimal solution. We will take a random guess and fix some of these AMs, i.e. we add new constraints to the MIP formulation fixing values of the respective variables to 1.

@@

First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set. This is because the specification of DTD/XSD does not allow an attribute to be \texttt{ID} and \texttt{IDREF} (\texttt{IDREFS}) at the same time. Let us denominate these mappings as \textit{IDREF candidates} (obviously different from \textit{candidate AMs}).
